---
title: "Palmer Penguins Report, exercise 2"
output: html_document
date: "2023-08-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readr)
penguins <- read_csv("data/penguins.csv")
```

In this report, we showcase the relationship between measurments taken of various penguin species in Antarctica. 

## Exploring the data

The `penguins` dataset was originally sourced from observations collected at the Palmer Station in Antarctica. The dataset is shown below. It contains **`r nrow(penguins)`** rows.

```{r}
penguins
```

### Inspecting missing values

At first glance, it appears that the dataset has some variables with missing values. Below we use the code chunk to explore which observations lack measurement for `body_mass_g`. Are there any other variables with missing values? 

```{r missing-values}
filter(penguins, is.na(body_mass_g))
```

## Visualization

The best way to examine the data is through visualizing the relationship between its variables. The plots below tell us a bit more about the penguin species in the dataset. 

## Penguin bills

Let's examine the data with a plot. We will focus here on the relationship between bill length and bill depth. 

```{r}
library(ggplot2)
ggplot(
  data = species_bills,
  mapping = aes(x = bill_length_mm, y = bill_depth_mm, color = species)
) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  theme_minimal() +
  labs(
    title = "Within species, longer bills are deeper",
    x = "Bill length (mm)",
    y = "Bill depth (mm)"
  )
```

We display the distribution of the penguin bill length below.

```{r plot1, message=FALSE, warning=FALSE}
ggplot(
  data = species_bills,
  mapping = aes(x = bill_length_mm)
) +
  geom_histogram() + 
  theme_minimal() +
  labs(
    x = "Bill length (mm)"
  )
```
