---
title: "Session 5"
date: September 5, 2024
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Agenda

* dplyr continued
* Create new variables with `mutate()`
* Group data using `group_by()`
* Create (grouped) summaries with `summarise()` 


## Script

You can download the session 5 script by clicking the link below.

<!-- to add scripts -->


<!-- ```{r, include = FALSE, eval = FALSE} -->

<!-- # install and load these packages -->
<!-- devtools::install_github('yihui/xfun') -->

<!-- # embed_file() requires a few more packages -->
<!-- xfun::pkg_load2(c('base64enc', 'htmltools', 'mime')) -->
<!-- ``` -->



<!-- ```{r echo=FALSE} -->
<!-- # a embed single file -->
<!-- xfun::embed_file('session_05.R') -->
<!-- ``` -->




## Resources

We will continue to talk about things covered in Chapter 5, this time sections 5.5 - 5.7. 

The `kableExtra` package is great for styling tables created by for example `summarise()`. Information about this package can be found [here](https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html).


## Homework Exercises

The session 5 homework is below. All code solutions should be completed using **one** (sometimes rather long) pipeline. 

```{r, include = FALSE}
library(tidyverse)
library(readxl)
library(kableExtra)

cb_data <- read_xlsx(path = "CB_data_2023.xlsx",
                     skip = 11,
                     na = "-99")

```

1. Recreate the code to make the plot below. Make sure your code includes the fix we talked about during the Thursday session to correct the Weight observations made by `raID-07`.

    ```{r message=FALSE, warning=FALSE, echo=FALSE, fig.width=5, fig.align="center", fig.height=3.6}
    cb_data %>% 
      mutate(Weight_fixed = ifelse(Observer == "raID-07",
                                   Weight * 1000,
                                   Weight),
             Ws = sqrt(Weight_fixed)) %>% 
      ggplot(aes(x = Ws,
                 y = Volume)) +
      geom_point() +
      theme_minimal() +
      xlab(expression(sqrt(Weight)))

    ```

1. Use `group_by()` and `mutate()` to calculate the median Volume per Observer, and call the new column `median_Volume`.

1. Find the largest Volume measurement per phase, using `group_by()`, `summarise()` and `max()`. 

1. How many observations per phase has a Weight less than 10 grams? 

1. Mean-center (subtract the mean from all the individual observations) the `Weight` variable (corrected). What is the mean of this new variable?

1. Mean-center the Volume variable for each phase separately.

1. Recreate the plot below, but instead of asking `geom_bar()` and `geom_errorbar()` to calculate the mean and the standard error, calculate these values yourself using `group_by()` and `summarise()`. Once these summary statistics are calculated, pipe to `ggplot()` and add the code necessary to produce the plot. Use google to figure out how to calculate standard errors and how to use pre-calculated values with `geom_bar()` and `geom_errorbar()`.  

    ```{r message=FALSE, warning=FALSE, echo=FALSE, fig.width=5, fig.align="center", fig.height=3.6}
    cb_data %>% 
      group_by(`Phase (color)`) %>% 
    summarise(mean = mean(Volume, na.rm = TRUE),
              sd = sd(Volume, na.rm = TRUE),
              n = n(),
              sterr = sd/sqrt(n)) %>% 
  ggplot(aes(x = `Phase (color)`,
             y = mean, 
             fill = `Phase (color)`)) +
  geom_bar(stat = "identity", 
           color = "black",
           show.legend = FALSE) +
  geom_errorbar(aes(ymin = mean-sterr, 
                    ymax = mean+sterr),
                width = 0.3) +
  ylab("Mean Volume (+/- Standard Error)") +
      theme_minimal() +
       scale_fill_manual(values = c("1-lightgray" = "lightgray",
                                "2-darkgray" = "darkgray",
                                "3-darkorange" = "darkorange",
                                "4-orange" = "orange",
                                "5-orangered" = "orangered",
                                "6-red" = "red")) 

    ```

1. Using the plot above as a starting point, create two new plots; one with errorbars representing the median absolute deviation and one with errorbars representing 95% confidence intervals (Hint: 95% CIs are calculated by multiplying the standard error with a certain constant). 

1. During session 4, we talked about how to use the `DT` package to create interactive tables in R markdown. These are perhaps most suited for large tables where showing all information would be problematic without the ability to produced paged, clickable tables. Many times when working in R we produce smaller tables, which is often the case when working `group_by()` and `summarise()` for example. Out of the box, these tables don't look very good when knitting a html document. Check the resources and see if you can reproduce the code necessary to make the table below. 


    ```{css, echo=FALSE}
table tr td p {
  margin: 4px;
}

thead p {
  margin: 4px;
}

caption p {
  margin: 0px;
}

    ```

    ```{r, echo=FALSE}
cb_data %>% 
      group_by(`Phase (color)`) %>% 
      summarise(mean = round(mean(Volume, na.rm = TRUE), 2),
                stdev = round(sd(Volume, na.rm = TRUE), 2),
                n = n()) %>% 
      kbl(caption = "Table 1: Cloudbuddy Volume") %>% 
      kable_classic(full_width = F) %>% 
      row_spec(1, color = "black", background = "lightgray") %>% 
      row_spec(2, color = "black", background = "darkgray") %>% 
      row_spec(3, color = "black", background = "darkorange") %>% 
      row_spec(4, color = "black", background = "orange") %>% 
      row_spec(5, color = "black", background = "orangered") %>% 
      row_spec(6, color = "black", background = "red") 
      
      
    ```

1. Calculate the difference in mean Volume between observations younger than the median age and observations older than the median age. The `diff()` function might be useful in this case. 

1. Calculate the minimum Volume for each cloudbuddy and phase. Next, calculate the mean of all the minimum values you just generated (the output should contain one value only). You might want to check out chapter 5.6.6 in the R4DS text book to figure out how to get things to run properly. Does it make a difference if you remove NAs in the beginning of the pipeline using `filter()`, if you use `na.rm = TRUE` when calculating the minimum values, or if you use `na.rm = TRUE` when calculating the mean? Pay attention to messages produced (warnings and such) when trying the three different options. Explain any differences you see in the output. 





 

   
